### 2020-07-01 manuscript.tex
Update
fixes

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Smarter dating moves and faster proposals: revisiting the phylogenetic relaxed clock model} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Jordan Douglas\textsuperscript{1*},
Rong Zhang\textsuperscript{1},
Alexei J. Drummond\textsuperscript{1,2},
Remco Bouckaert\textsuperscript{1}
\\
\bigskip
\textbf{1} Centre for Computational Evolution, School of Computer Science, University of Auckland, Auckland, New Zealand\\
\textbf{2} School of Biological Sciences, University of Auckland, Auckland, New Zealand
\\
\bigskip

@@

\subsection*{A guided adaptive leaf rate operator}

@@

\subsubsection*{AVMVN kernel}
\label{AVMVN_sect}

@@

\subsection*{Clock model averaging}
\label{sect:clockModelAveraging}

In Bayesian model averaging (BMA), competing models are estimated in the posterior distribution alongside other parameters and are thus directly selected through usual Bayesian inference methods  \cite{fragoso2018bayesian}. This is in contrast to other methods -- such as path sampling \cite{lartillot2006computing} and stepping-stone sampling \cite{xie2011improving} -- which operate outside of this central Bayesian inference process (in this case the MCMC algorithm). BMA has been applied to many domains \cite{fragoso2018bayesian} -- including astrophysics \cite{parkinson2013bayesian}, geographical information systems \cite{barber2006modelling}, and molecular biology \cite{douglas2020bayesian} -- and in phylogenetics  has been applied to selection of substitution models \cite{beier2004phylogenetic,posada2008jmodeltest,bouckaert2017bmodeltest}, rate heterogeneity models \cite{bouckaert2017bmodeltest}, and clock models \cite{li2012model,baele2012accurate}.

Li and Drummond 2012 used Bayesian model averaging to select clock model prior distributions -- LogNormal, Exponential, and Inverse Gaussian \cite{li2012model}. This was achieved by a model indicator that points to the applicable clock model. However, due to its discrete nature, sampling the model indicator can be inefficient.

As a preliminary non-phylogenetic analysis, we explored the statistical power of distinguishing between log-normally and exponentially distributed data, and how this depends on sample size $2N-2$ (for $N$ taxa) and standard deviation $\sigma$ (for log-normally distributed data). Models are distinguished using the Bayesian information criterion (BIC):

\begin{align}
	BIC = \begin{cases} \log(2N-2) - 2 \log \big( \hat{L} \; P(\hat{\sigma}) \big) & \text{for LogNormal}(-0.5\sigma^2, \sigma) \\ 
	- 2 \log \big( \hat{L} \big) & \text{for Exponential}(1) \end{cases}
\end{align}

where $\hat{L}$ is the maximim likelihood and $P(\hat{\sigma})$ is the prior density of maximum likelihood estimate $\hat{\sigma}$. This analysis reveals that when there are a small number of rates, data simulated from a lognormal distribution is often best modelled with an exponential distribution, especially when $\sigma$ is close to 1 (\text{Fig. \ref{fig:ClockModelAverging}}). However, for larger trees $N > 30$, clock model averaging should be a viable approach.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{Figures/ClockModelAverging.pdf}
\caption{\textbf{Ability to resolve LogNormal and Exponential distributed data.} \textit{Left:} at each grid point, $2N-2$ random variables are sampled from a $\text{LogNormal}(-0.5\sigma^2, \sigma)$ distribution. The samples are assessed for their fit to i) a $\text{LogNormal}(-0.5\hat{\sigma}^2, \hat{\sigma})$ distributon, and ii) an $\text{Exponential}(1)$ disribution, using the Bayesian information criterion (BIC). This is repeated 1000 $\times$ at each grid point and the two sets of BIC scores are compared using a paired t-test ($p=0.001$), which determines the colour of the point on the plot. \textit{Right:} points are instead sampled from an $\text{Exponential}(1)$ disribution. These simulations show that clock model averaging is most applicable for trees with more than 30 taxa.   } 
\label{fig:ClockModelAverging}
\end{figure}

@@

1  &  6  &  16  &  13.7  &  1.1  &  Spiders (Richart 2015 \cite{Richart_2016}) \\

2  &  18  &  16  &  6.7  &  0.9  &  Gallopheasants (Meiklejohn 2016 \cite{Meiklejohn_2016}) \\

3  &  33  &  16  &  5.1  &  1.5  &  Birds (McCormack 2013 \cite{McCormack_2013}) \\

4  &  40  &  1 &  3.0  &  0.6-1.8  &  \textit{Simulated data} $\times 10$ \\

5  &  44  &  3  &  1.9  &  0.8  &  Bark beetles (Cognato 2001 \cite{Cognato_2001}) \\

6  &  51  &  6  &  5.4  &  1.8  &  Southern beeches (Sauquet 2011 \cite{Sauquet_2011}) \\

7  &  70  &  3  &  2.2  &  0.9  &  Caterpillars (Kawahara 2013 \cite{Kawahara_2013}) \\

8  &  94  &  4  &  1.8  &  0.8  &  Bees (Rightmyer 2013 \cite{Rightmyer_2013}) \\

9  &  106  &  1  &  0.6  &  0.3  & Songbirds (Moyle 2016 \cite{Moyle_2016}) \\

10  &  110  &  1  &  0.5  &  0.3  &  Vertebrates (Fong 2012 \cite{Fong_2012}) \\

 \hline
\end{tabular}
\caption{Datasets used during benchmarking, sorted in increasing order of taxa count $N$. Number of partitions $P$, total alignment length $L$, and number of patterns $L_\text{eff}$ are also specified. The 9 empirical datasets are benchmarked $10 \times$ per setting, whereas each of the 10 simulated datasets are benchmarked once per setting.}
\label{table:datasets}
\end{table}

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Smarter dating moves and faster proposals: revisiting the phylogenetic relaxed clock model} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Jordan Douglas\textsuperscript{1,2*},
Rong Zhang\textsuperscript{1,2},
Alexei J. Drummond\textsuperscript{1,2,3},
Remco Bouckaert\textsuperscript{1,2}
\\
\bigskip
\textbf{1} Centre for Computational Evolution,  University of Auckland, Auckland, New Zealand\\
\textbf{2} School of Computer Science, University of Auckland, Auckland, New Zealand\\
\textbf{3} School of Biological Sciences, University of Auckland, Auckland, New Zealand
\\
\bigskip

@@

\subsection*{A guided adaptive leaf rate operator}
\label{AVMVN_sect}

@@

\subsubsection*{AVMVN kernel}

@@

@@

 1  &  38  &  8  &  9.1  &  6.45  &  Seed plants (Ran 2018 \cite{Ran_2018}) \\

2  &  44  &  7  &  5.9  &  1.8  &  Squirrel Fishes (Dornburg 2012 \cite{Dornburg_2012}) \\

3  &  44  &  3  &  1.9  &  0.8  &  Bark beetles (Cognato 2001 \cite{Cognato_2001}) \\

4  &  51  &  6  &  5.4  &  1.8  &  Southern beeches (Sauquet 2011 \cite{Sauquet_2011}) \\

5  &  61  &  8  &  6.9  &  4.3  &  Bony fishes (Broughton 2013 \cite{Broughton_2013}) \\

6  &  70  &  3  &  2.2  &  0.9  &  Caterpillars (Kawahara 2013 \cite{Kawahara_2013}) \\

7  &  78  &  8  &  3.4  &  3.1  &  Animals (Cannon 2016 \cite{Cannon_2016}) \\

8  &  80  &  1 &  10.0  &  4.2  &  \textit{Simulated data}  \\

9  &  94  &  4  &  2.2  &  1  &  Bees (Rightmyer 2013 \cite{Rightmyer_2013}) \\

10  &  106  &  1  &  0.8  &  0.5  &  Songbirds (Moyle 2016 \cite{Moyle_2016}) \\

 \hline
\end{tabular}
\caption{Datasets used during benchmarking, sorted in increasing order of taxa count $N$. Number of partitions $P$, total alignment length $L$, and number of patterns $L_\text{eff}$ are also specified.}
\label{table:datasets}
\end{table}
